Making Sequential Programs Deterministic

### 2019-04-05 experiments.tex
Edits in DeFlaker and threats to validity

\section{Comparison to Deflaker}
Deflaker~\cite{deflaker}, like NonDex, reports tests that could be flaky. 
We were unable to run DeFlaker on Randoop, because Deflaker works with projects built with Maven and Randoop
uses Gradle for its build system. We plan to convert Randoop's build system to Maven and
perform this case study.
%Ideally, we would have 
%liked to run all three tools (NonDex, DeFlaker, and \TheDeterminismChecker) on an open source project
%that wasn't explored by any of the tool for a fair comparison. Unfortunately, there are issues with the
%DeFlaker plugin and we weren't able to run it on any project. We have created a Github issue for the same
%in its repository.
%\todo{Deflaker works now. Tried on retrofit: only works where there is sufficient change in code coverage between head and its parent.
%Looked throught= about 10 commits, didn't find differential code coverage and therefore no flakes}
%\todo{Deflaker requires Maven build. Randoop is on Gradle.}

\subsection{\TheDeterminismChecker on DeFlaker benchmarks}\label{sec:deflaker-benchmarks}
DeFlaker found 87 previously unknown flaky tests in 93 projects that were being actively developed at the the time
the paper was written. The authors reported 19 of these tests, out of which 7 were addressed by the maintainers of
those projects. Since we can only verify if \TheDeterminismChecker finds the same bugs by looking at the source code
locations, we ran our tool on the benchmarks where the reported bug was fixed. We again defer the reader to~\cite{deflaker}
for a complete list of the open source projects and their details.

Two of the seven flaky tests (in \<achilles> and \<jackrabbit-oak>) that DeFlaker reported were caused by a race condition. \TheDeterminismChecker
is designed for sequential programs and hence cannot detect this category of bugs.  
We couldn't find the link to one of the issues in the issue tracker of \<jackrabbit-oak>
We ran into build issues
for \<tigglz> and \<nutz> which had two and one flakiness issues respectively. 
\TheDeterminismChecker was able to find the source of flakiness in \<checkstyle>. This bug was in a test
case that deterministically accessed an \<OrderNonDet> array returned by \<Class.getDeclaredConstructors()>.

>>>>

\section{Comparison to Deflaker}
Deflaker~\cite{deflaker}, like NonDex, reports tests that could be flaky. 
We were unable to run DeFlaker on Randoop, because Deflaker works with
projects built with Maven, but Randoop
uses Gradle for its build system. We plan to convert Randoop's build system to Maven and
perform this case study.
%Ideally, we would have 
%liked to run all three tools (NonDex, DeFlaker, and \TheDeterminismChecker) on an open source project
%that wasn't explored by any of the tool for a fair comparison. Unfortunately, there are issues with the
%DeFlaker plugin and we weren't able to run it on any project. We have created a Github issue for the same
%in its repository.
%\todo{Deflaker works now. Tried on retrofit: only works where there is sufficient change in code coverage between head and its parent.
%Looked throught= about 10 commits, didn't find differential code coverage and therefore no flakes}
%\todo{Deflaker requires Maven build. Randoop is on Gradle.}

\subsection{\TheDeterminismChecker on DeFlaker benchmarks}\label{sec:deflaker-benchmarks}
DeFlaker found 87 previously unknown flaky tests in 93 projects that were being actively developed at the the time
the paper was written. The authors reported 19 of these tests, out of which 7 were addressed by the maintainers of
those projects~\cite{deflaker}.
We ran \theDeterminismChecker on the part of each of these
codebases where the reported bug was fixed, as in \cref{sec:nondex-benchmarks}.
% The DeFlaker authors graciously fixed problems we discovered while using
% their tool.

Two of the seven flaky tests (in \<achilles> and \<jackrabbit-oak>)
were caused by a race condition, which \theDeterminismChecker
cannot detect.  (This is a strength of DeFlaker over \theDeterminismChecker.)
We couldn't find the link to one of the issues in the issue tracker of \<jackrabbit-oak>.
We were unable to build \<tigglz> and \<nutz> which had two and one flakiness issues respectively.
\TheDeterminismChecker was able to find the source of flakiness in \<checkstyle>. This bug was in a test
case that treated an array returned by \<Class.getDeclaredConstructors()>
as deterministic.

### 2019-04-05 threats.tex
taint citations
Edits in DeFlaker and threats to validity

The type system presented in this paper does not account for the fact that
a nondeterministic conditional expression (in an \<if> or \<switch>
statement or a \<for> or \<while> loop, for example) may behave differently
from execution to execution.
These ``implicit flows'' are a well-known challenge for taint analysis.
Suppose that an expression used in a conditional is tainted.
A standard approach\todo{Add citations} for a dynamic taint analysis is to taint any value that may
be assigned in either branch of the conditional statement.  This tends to rapidly lead to most of
the program state being tainted.
In a static analysis, every possibly-assigned lvalue would be considered
possibly tainted; again, this tends to snowball through the program.

@@

It our case study, we also disabled a check that required all caught
exceptions to have \<NonDet> type.  This type is a safe approximation to an
exception that can be thrown from arbitrary code, including libraries.
Enabling this check led to many unhelpful false positive warnings.

\TheDeterminismChecker only examines the code it is run on.  Unchecked
libraries and native code with incorrect specifications might introduce
nondeterminism that it does not check.

We have performed only one case study.  This case study found important
previously-unknown errors, but its results might not carry over to other
programs or to programs written in other programming lanugages.  We
mitigated this problem by showing that \theDeterminismChecker finds most or
all of the sources of nondeterminism identified by other tools.

>>>>

The type system presented in this paper does not account for the fact that
a nondeterministic conditional expression (in an \<if> or \<switch>
statement or a \<for> or \<while> loop, for example) may behave differently
from execution to execution.
These ``implicit flows'' are a well-known challenge for taint analysis.
Suppose that an expression used in a conditional is tainted.
A standard approach
\cite{Kang2011DTADT,Arzt:2014:FPC:2594291.2594299}
%\todo{Add citations}
 for a dynamic taint analysis is to taint any value that may
be assigned in either branch of the conditional statement.  This tends to rapidly lead to most of
the program state being tainted.
In a static analysis, every possibly-assigned lvalue would be considered
possibly tainted; again, this tends to snowball through the program.

@@

It our case study, we also disabled a check that required all caught
exceptions to have \<NonDet> type.  This type is a safe approximation to an
exception that can be thrown from arbitrary code, including libraries.
Enabling this check led to many unhelpful false positive warnings, which
are only relevant if unverified libraries throw nondeterministic exceptions.

\TheDeterminismChecker only examines the code it is run on.  Unchecked
libraries and native code with incorrect specifications might introduce
nondeterminism even if \theDeterminismChecker issues no warnings.

We have performed only one case study.  This case study found important
previously-unknown errors, but its results might not carry over to other
programs or to programs written in other programming lanugages.  We
mitigated this problem by showing that \theDeterminismChecker finds many
of the sources of nondeterminism identified by other tools.
